# Exploring methods for sex ratios in age and length composition data {#sexratios}

One of the needed improvements is dealing with sex ratios in the compositional data. Currently it is difficult to see how there is any information on sex ratio in the Alaskan sablefish stock assessment given how the observations are structured (Chapter \@ref(modeldescription)) (Caveat is users can provide an observed sex ratio for observations etc which is similar to the approach in @ward2019assessing). I am aware of two approaches for supplying observations that in theory should provide information on sex ratio. The first (approach 1) was taken from Casal2 [@doonan2016casal2]. This treats sexed composition data for a year as a single proportion i.e., proportions across all ages and sexes sum to one for each year,

\[
\boldsymbol{P}^k_{y} = \frac{(C^k_{a,y,1},C^k_{a,y,2})}{\sum_a \sum_s C^k_{a,y,s}}, \quad \sum \boldsymbol{P}^k_{y} = 1
\]
where, \(C^k_{a,y,1}\) is the catch at age (numbers) for males in year \(y\), age \(a\) and survey \(k\). \(\boldsymbol{P}^k_{y}\) is a proportion vector that sums to one that covers both sexes and corresponding ages. The likelihood contribution for this approach follows  
\[
\boldsymbol{P}^k_{y} \sim Multinomial(\mathbb{E}[\boldsymbol{P}^k_{y}], N^{eff, k}_{y}) \ .
\]
where, \(N^{eff, k}_{y}\) is the effective sample size for this survey and year.




The second approach (approach 2) is to treat composition for each sex seperately that is proportions at age or length for a sex will sum to one, but also provide a specific sex ratio observation over all ages or lengths as done in the New Zealand rock lobster stock assessment [@webber2021rapid]. 

\[
R^k_{y,s} = \frac{\sum_a C^k_{a,y,s}}{\sum_a \sum_s C^k_{a,y,s}}, \quad  \sum_s R^k_{y,s} = 1
\]
and,
\[
P^k_{a, y,s} = \frac{C^k_{a,y,s}}{\sum_a C^k_{a,y,s}}, \quad  \sum_a P^k_{a, y,s} = 1 \ .
\]
The likelihood assumptions for this model are,
\[
R^k_{y,s} \sim Binomial(\mathbb{E}[R^k_{y,s}], \sum_s N^{eff, k}_{y,s})
\]
and,
\[
\boldsymbol{P}^k_{y,s} \sim Multinomial(\mathbb{E}[\boldsymbol{P}^k_{a, y,s}], N^{eff, k}_{y,s}) \ .
\]
where, \(N^{eff, k}_{y,s}\) is the effective sample size.



## Simulation testing {-}
A simple simulation experiment was conducted to explore the benefits of each approach. There were two OM scenarios that we wanted to explore, the first was that the true population had a 50:50 sex ratio and the observations had skewed sex ratios due to the selectivity i.e., availability to the survey fishery. Second the selectivity was constant for both sexes but the sex ratio was skewed at recruitment.



### Operating Model (OM) Parameters {-}

```{r setup}
bio_params = list(
  ages = 1:20,
  L_inf_m = 58,
  K_m = 0.133,
  t0_m = 0,
  L_inf_f = 62,
  K_f = 0.143,
  t0_f = 0,
  M = 0.15,
  a_m = 2.18e-9, ## tonnes
  b_m = 3.2,
  a_f = 2.08e-9, ## tonnes
  b_f = 3.3,
  m_a50 = 2.3,
  m_ato95 = 1.2,
  sigma = 0.6,
  h = 0.85,
  sigma_r = 0.6,
  R0 = 8234132,
  plus_group = 1 # 0 = No, 1 = Yes
)

other_params = list(
  s_a50_m = 3.6,
  s_ato95_m = 2,
  s_a50_f = 4.6,
  s_ato95_f = 1.4,
  s_q = 0.2,
  s_alpha = 1,
  f_a50_m = 4.2,
  f_ato95_m = 1.17,
  f_a50_f = 4.7,
  f_ato95_f = 1.76,
  f_alpha = 1,
  ssb_prop_Z = 0.5,
  survey_prop_Z = 0.5,
  survey_age_error = c(0.5, 0.4),  ## sd, rho (ignored if iid)
  fishery_age_error = c(0.5, 0.4),  ## sd, rho (ignored if iid)
  survey_bio_cv = c(0.1)
)

ages = bio_params$ages
max_age = max(bio_params$ages)
n_years = 30
years = (2020 - n_years + 1):2020
n_ages = length(ages)
## annual fishing mortality
start_F = c(rlnorm(10, log(seq(from = 0.02, to = 0.1, length = 10)), 0.1), rlnorm(10, log(0.1), 0.1), rlnorm(10, log(0.07), 0.1))
recruit_devs = log(rlnorm(n_years, -0.5 * bio_params$sigma_r * bio_params$sigma_r, bio_params$sigma_r))
```


```{r Build_tmp_sex, echo = F, eval = T}
##################
## Plot Biology
##################
length_at_age_m = vonbert(bio_params$ages, bio_params$K_m, bio_params$L_inf_m, bio_params$t0_m)
length_at_age_f = vonbert(bio_params$ages, bio_params$K_f, bio_params$L_inf_f, bio_params$t0_f)
fishery_ogive_m = logis(bio_params$ages, other_params$f_a50_m, other_params$f_ato95_m)
fishery_ogive_f = logis(bio_params$ages, other_params$f_a50_f, other_params$f_ato95_f)

survey_ogive_m = logis(bio_params$ages, other_params$s_a50_m, other_params$s_ato95_m)
survey_ogive_f = logis(bio_params$ages, other_params$s_a50_f, other_params$s_ato95_f)
mat_age = logis(bio_params$ages, bio_params$m_a50, bio_params$m_ato95)
weight_at_age_m = bio_params$a_m * length_at_age_m^bio_params$b_m
weight_at_age_f = bio_params$a_m * length_at_age_f^bio_params$b_f

## can change these if we want biennial surveys, or truncated observatons.
## Using the ALN method doesn't allow zeros so consider truncating tail ages.
survey_year_obs = seq(from = min(years), to = max(years), by = 2)
survey_ages = ages
fishery_year_obs = seq(from = min(years) + 1, to = max(years), by = 2)
fishery_ages = ages

############
## Build a multinomial model to double check estimability of all parameters
## In this case we have 'good' data, annual data, no ageing error.
##  survey index cv = 0.05
##  year effective sample size = 1000
############
TMB_data = list()
TMB_data$ages = ages
TMB_data$maxAgePlusGroup = 1;
TMB_data$years = years
TMB_data$nyears = length(TMB_data$years)
TMB_data$nages = length(TMB_data$ages)

## No ageing error
TMB_data$ageing_error_matrix = matrix(0, nrow = TMB_data$nages, ncol = TMB_data$nages)
diag(TMB_data$ageing_error_matrix) = 1;

TMB_data$survey_year_indicator = as.integer(TMB_data$years %in% survey_year_obs)
TMB_data$survey_obs = rnorm(sum(TMB_data$survey_year_indicator), 100, 4)
TMB_data$survey_cv = rep(0.15,sum(TMB_data$survey_year_indicator))
TMB_data$survey_sample_time = rep(0.5,sum(TMB_data$survey_year_indicator))
TMB_data$survey_AF_obs = array(2, dim = c(n_ages * 2, sum(TMB_data$survey_year_indicator)))
TMB_data$survey_AF_type = 0;
TMB_data$survey_numbers_male = colSums(TMB_data$survey_AF_obs)/ 2
TMB_data$fishery_year_indicator = as.integer(TMB_data$years %in% fishery_year_obs)
TMB_data$fishery_AF_obs = array(2, dim = c(n_ages * 2, sum(TMB_data$fishery_year_indicator)))
TMB_data$fishery_AF_type = 0;
TMB_data$fishery_numbers_male = colSums(TMB_data$fishery_AF_obs)/ 2
TMB_data$catches = rep(1000, n_years)# this will be overriden in the simulate() call
TMB_data$ycs_estimated = rep(1, n_years)
TMB_data$standardise_ycs = 0;

TMB_data$catchMeanLength = TMB_data$stockMeanLength = array(0, dim = c(TMB_data$nages, TMB_data$nyears,2))
TMB_data$catchMeanLength[,,1] = TMB_data$stockMeanLength[,,1] = matrix(length_at_age_m, byrow = F, ncol = TMB_data$nyears, nrow = TMB_data$nages)
TMB_data$catchMeanLength[,,2] = TMB_data$stockMeanLength[,,2] =  matrix(length_at_age_f, byrow = F, ncol = TMB_data$nyears, nrow = TMB_data$nages)

TMB_data$prop_mature = matrix(mat_age, byrow = F, ncol = TMB_data$nyears, nrow = TMB_data$nages)
TMB_data$natMor = bio_params$M
TMB_data$steepness = bio_params$h
TMB_data$stockRecruitmentModelCode = 2 ## BH
TMB_data$propZ_ssb = rep(other_params$ssb_prop_Z, TMB_data$nyears)
TMB_data$propZ_survey = rep(other_params$survey_prop_Z, TMB_data$nyears)
TMB_data$sel_ato95_bounds = c(0.1,20)
TMB_data$sel_a50_bounds = c(0.1,20)
TMB_data$sel_alpha_bounds = c(0.5, 2)

TMB_data$mean_weight_a = c(bio_params$a_m,bio_params$a_f) 
TMB_data$mean_weight_b = c(bio_params$b_m,bio_params$b_f)


## The same parameters as OM, to check for consistency
true_pars = list(
  ln_R0 = log(bio_params$R0),
  ln_ycs_est =  log(exp(recruit_devs[TMB_data$ycs_estimated == 1] - 0.5*bio_params$sigma_r^2)),
  ln_sigma_r = log( bio_params$sigma_r),
  ln_extra_survey_cv = log(0.0001),
  logit_f_a50 = logit_general(c(other_params$f_a50_m, other_params$f_a50_f), TMB_data$sel_a50_bounds[1], TMB_data$sel_a50_bounds[2]),
  logit_f_ato95 = logit_general(c(other_params$f_ato95_m, other_params$f_ato95_f), TMB_data$sel_ato95_bounds[1], TMB_data$sel_ato95_bounds[2]),
  logit_f_alpha_f = logit_general(other_params$f_alpha, TMB_data$sel_alpha_bounds[1], TMB_data$sel_alpha_bounds[2]),
  logit_survey_a50 = logit_general(c(other_params$s_a50_m, other_params$s_a50_f), TMB_data$sel_a50_bounds[1], TMB_data$sel_a50_bounds[2]),
  logit_survey_ato95 = logit_general(c(other_params$s_ato95_m, other_params$s_ato95_f), TMB_data$sel_ato95_bounds[1], TMB_data$sel_ato95_bounds[2]),
  logit_surveyQ = qlogis(other_params$s_q),
  logit_survey_alpha_f = logit_general(other_params$s_alpha, TMB_data$sel_alpha_bounds[1], TMB_data$sel_alpha_bounds[2]),
  ln_F = log(start_F),
  logit_proportion_male = rep(0, TMB_data$nyears),
  ln_catch_sd = log(0.02)
)
##################################
### Build TMB OM with Multinomial
##################################
#dyn.unload(dynlib(file.path(DIR$book,"TMB","SexDisaggregatedAgeStructuredModel")))
compile(file.path("TMB","SexDisaggregatedAgeStructuredModel.cpp"), flags = "-Wignored-attributes -O3",DLLFLAGS="");
dyn.load(dynlib(file.path("TMB","SexDisaggregatedAgeStructuredModel")))
## tolerance form model convergence, all gradients need to be less than this.
grad_tol = 0.001
# these parameters we are not estimating.
na_map = fix_pars(par_list = true_pars, pars_to_exclude = c("ln_catch_sd","ln_extra_survey_cv","ln_sigma_r", "logit_proportion_male"))
ASM_obj <- MakeADFun(TMB_data, true_pars, DLL= "SexDisaggregatedAgeStructuredModel", map = na_map,checkParameterOrder=FALSE)
TMB_data_alt = TMB_data
TMB_data_alt$survey_AF_type = 1
TMB_data_alt$fishery_AF_type = 1
ASM_obj_alt_sex <- MakeADFun(TMB_data_alt, true_pars, DLL= "SexDisaggregatedAgeStructuredModel", map = na_map,checkParameterOrder=FALSE)

##################################
### Generate predictions from both methods to compare
##################################
sim_data = ASM_obj$simulate(complete = T)
sim_data_alt = ASM_obj_alt_sex$simulate(complete = T)
```




```{r compile_sex_plot, warning=F}
par(mfrow = c(2,2))
plot(survey_ages, sim_data_alt$survey_comp_fitted[1:n_ages,9], type = "l", lwd = 3, ylab = "Survey predicted values", main = "Approach 1", xlab = "", ylim = c(0,0.2))
lines(survey_ages, sim_data_alt$survey_comp_fitted[(n_ages + 1):(n_ages*2),9], type = "l", lwd = 3, lty = 2, col = "red")
plot(survey_ages, sim_data$survey_comp_fitted[1:n_ages,9], type = "l", lwd = 3, ylab = "", main = "Approach 2", xlab = "", ylim = c(0,0.2))
lines(survey_ages, sim_data$survey_comp_fitted[(n_ages + 1):(n_ages*2),9], type = "l", lwd = 3, lty = 2, col = "red")

plot(fishery_ages, sim_data_alt$fishery_comp_fitted[1:n_ages,9], type = "l", lwd = 3, ylab = "Fishery predicted values", main = "", xlab = "", ylim = c(0,0.3))
lines(fishery_ages, sim_data_alt$fishery_comp_fitted[(n_ages + 1):(n_ages*2),9], type = "l", lwd = 3, lty = 2, col = "red")

plot(fishery_ages, sim_data$fishery_comp_fitted[1:n_ages,9], type = "l", lwd = 3, ylab = "", main = "", xlab = "", ylim = c(0,0.3))
lines(fishery_ages, sim_data$fishery_comp_fitted[(n_ages + 1):(n_ages*2),9], type = "l", lwd = 3, lty = 2, col = "red")
legend('topright', legend = c("Male","Female"), col = c("black", "red"), lty = c(1,2), lwd = 3)
```


### Scenario 1 {-}
The first simulation assumed the OM had 50:50 males and female sex ratio at recruitment, and selectivities were as described in the above OM section. We simulated 100 data sets for each of the two approach's for sexually disaggregated compositional data outlined in the introduction. These were then fitted to in two EM's, where the EM's differed in the approach they used. Both EM's estimated a scalar on the female selectivity that allowed females to be more or less selected compared to males for both the fishery and survey. This was done be introducing an estimable \(\alpha\) parameter into the selectivity. Males selectivities were constrained to have a max value = 1 using,
\[
S^{male}_a = 1/(1+19^{(a_{50}-a)/a_{to95})}) 
\]
where as the female was
\[
S^{female}_a = \alpha/(1+19^{(a_{50}-a)/a_{to95})}) 
\]

Although this scenario assumed both male and female had the same max selectivity I wanted to see the effect of estimating this additional parameter.

```{r run_scenario_1, warning=F, echo = F, eval = T}
##################################
### Simulate 100 data sets for Scenario 1
##################################
n_sims = 100
est_pars = NULL
ssbs = NULL
convergence = NULL
max_grad = NULL
est_list = list()
for(sim_iter in 1:n_sims) {
  #if(sim_iter %% 10 == 0)
  #  cat("iter = ", sim_iter, "\n")
  sim_data = ASM_obj$simulate(complete = T)
  start_pars = ran_start_vals()
  est_model = MakeADFun(sim_data, start_pars, DLL= "SexDisaggregatedAgeStructuredModel", map = na_map, silent = T)
  opt_modelc = nlminb(est_model$par, est_model$fn, est_model$gr, control = list(iter.max = 10000, eval.max = 10000))
  ## Do two newton steps to try get parameters closer to the solution
  for(i in 1:2) {
    g = as.numeric(est_model$gr(opt_modelc$par))
    h = optimHess(opt_modelc$par, fn = est_model$fn, gr = est_model$gr)
    opt_modelc$par = opt_modelc$par - solve(h,g)
    opt_modelc$objective = est_model$fn(opt_modelc$par)
  }
  convergence[sim_iter] = opt_modelc$convergence
  max_grad[sim_iter] = max(abs(est_model$gr(est_model$env$last.par.best)))
  est_pars = rbind(est_pars, opt_modelc$par)
  est_rep = est_model$report(opt_modelc$par)
  ssbs = rbind(ssbs, est_rep$ssb)
  est_list[[sim_iter]] = est_rep
}
cat("max gradient from 100 simualtions = ", max(max_grad), "\n")
R0s = get_list_obj(est_ls = est_list, object_label = "R0")
sel_survey = get_list_obj(est_ls = est_list, object_label = "survey_selectivity")
sel_survey_df = data.frame(sim = factor(sel_survey[,1]), male = sel_survey[,2], female = sel_survey[,3], age = ages)
sel_fishery = get_list_obj(est_ls = est_list, object_label = "fishery_selectivity")
sel_fishery_df = data.frame(sim = factor(sel_fishery[,1]), male = sel_fishery[,2], female = sel_fishery[,3], age = ages)

## plot selectivities
male_s_sel = ggplot(data  = sel_survey_df) +
  geom_line(aes(x = age, y = male, group = factor(sim), alpha = 0.3, col = "EMs"), linewidth = 1.3) +
  geom_line(data = data.frame(age = ages, male = true_report$survey_selectivity[,1]), aes(x = age, y = male, col = "OM"), linewidth= 1.2, linetype = "dashed")+
  guides(alpha = "none") +
  labs(x = "Age", y = "Selectivity", title = "Survey - Male", col = "")+
  theme_bw()

female_s_sel = ggplot(data  = sel_survey_df) +
  geom_line(aes(x = age, y = female, group = factor(sim), alpha = 0.3, col = "EMs"), linewidth = 1.3) +
  geom_line(data = data.frame(age = ages, female = true_report$survey_selectivity[,2]), aes(x = age, y = female, col = "OM"), linewidth= 1.2, linetype = "dashed")+
  guides(alpha = "none") +
  labs(x = "Age", y = "Selectivity", title = "Survey - Female", col = "")+
  theme_bw()
male_f_sel = ggplot(data  = sel_fishery_df) +
  geom_line(aes(x = age, y = male, group = factor(sim), alpha = 0.3, col = "EMs"), linewidth = 1.3) +
  geom_line(data = data.frame(age = ages, male = true_report$fishery_selectivity[,1]), aes(x = age, y = male, col = "OM"), linewidth= 1.2, linetype = "dashed")+
  guides(alpha = "none") +
  labs(x = "Age", y = "Selectivity", title = "Fishery - Male", col = "")+
  theme_bw()

female_f_sel = ggplot(data  = sel_fishery_df) +
  geom_line(aes(x = age, y = female, group = factor(sim), alpha = 0.3, col = "EMs"), linewidth = 1.3) +
  geom_line(data = data.frame(age = ages, female = true_report$fishery_selectivity[,2]), aes(x = age, y = female, col = "OM"), linewidth= 1.2, linetype = "dashed")+
  guides(alpha = "none") +
  labs(x = "Age", y = "Selectivity", title = "Fishery - Female", col = "")+
  theme_bw()

grid.arrange(grobs = list(male_s_sel, female_s_sel, male_f_sel, female_f_sel), ncol= 2)

## plpt SSBs
molten_ssbs = melt(ssbs)
colnames(molten_ssbs) = c("sim", "year", "value")
ggplot() +
  geom_line(data =molten_ssbs, aes(x = year, y = value, group = sim)) +
  geom_line(data = data.frame(year = 1:31, value = true_report$ssb),aes(x = year, y = value), col = "red", linetype = "dashed", linewidth = 1.2) +
  ylim(0,NA) +
  labs(x = "Time", y = "SSB (t)") +
  theme_bw()
```


## Other things to consider {-}
- LF observations with sexual dimorphism in growth. A possible reason LF's could be influencing assessment output is due to mis-specified growth ala 


