# How to calculate the plus group in models with markovian movement {#spatialInit}

In "standard" age-structured models the plus group is calculated using an approximation to the infinite geometric series i.e.,

Analytically solving the age-structure for the initial state of the partition. The partition consists of a vector of numbers at age for each category denoted by \(\textbf{N}\);
\[
\textbf{N} = (N_1, N_2, ... , N_{a_+})^T
\]

where \(N_{a_+}\) are the numbers in the plus group. The numbers at age for a closed population with density independent equilibrium state can be defined as,

\[
N_a = \left\{ \begin{array}{lcl}
	R_0 & \mbox{for} & a = 1
	\\ 
	R_0e^{-aM}  & \mbox{for} & 1 < a < a_+
    \\
	R_0e^{\sum_{i = a_+}^\infty-iM} & \mbox{for} & a = a_+
\end{array}\right.
\]

Initialisation for ages \(1 \leq a < a_+\) is easy, all you need to do is iterate the model \(a_+ - 1\) times and the age classes will be populated with there respected initial numbers at age. It is less simple for the plus group (\(N_{a_+}\)) because it can take many iterations to reach the initial state of this age class, denoted by the \(\infty\)

So, although the plus group may start at let say 30 years old it actually represents 30 and 31 and 32 up to some biological plausible value lets say 130, but mathematically can be thought of as \(\infty\). The plus group is modelled for practical reasons and care should be given when choosing the cut-off age. The plus group is an infinite geometric series that is defined by the common ratio \(r_{a_+}\).

\begin{align}
N_{a_+} &= R_0e^{\sum_{a = a_+}^\infty-aM}\\
        &= R_0 \frac{e^{M - a_+M}}{1 - e^{-M}}
\end{align}


```{r illustrateplusGroup}
M = 0.1
R0 = 4e6
plus_group_age = 30
N_age = R0 * exp(-M * 1:plus_group_age)
## numerically calculate it
N_plus_group =  tail(N_age, n = 1)
for(i in 1:10000) {
  N_plus_group = N_plus_group * exp(-M) + tail(N_age, n = 1) * exp(-M)
}
N_plus_group
## using the geometric series calculation
R0 * exp(-M - plus_group_age*M) / (1 - exp(-M))
```

The question becomes, how do we calculate this plus group when there is markovian movement as well as? To illustrate the methods that we want will extend the above R-code to have a region and age partition.


```{r spatialcase}
M = 0.1
R0 = 4e6
n_regions = 3
movement_matrix = matrix(0, nrow = n_regions, ncol = n_regions);
movement_matrix[1,] = c(0.7, 0.2, 0.1)
movement_matrix[2,] = c(0.1, 0.6, 0.3)
movement_matrix[3,] = c(0.15, 0.05, 0.8)
## rows are "from" cols are "to"
plus_group_age = 30
## partition
N_age = matrix(0, nrow = n_regions, ncol = plus_group_age)
## initialise with no movement initially
N_age[1,] = R0 * exp(-M * 1:plus_group_age) ##
N_age[2,] = R0 * exp(-M * 1:plus_group_age) ## 
N_age[3,] = R0 * exp(-M * 1:plus_group_age) ##

## apply an annual cycle 10000 times to see what the "initial conditions should be"
N_next_year_age = N_age
for(i in 1:1000) {
  ## recruitment
  N_next_year_age[,1] = R0 * exp(-M)
  ## ageing and mortality
  N_next_year_age[,2:plus_group_age] = N_age[,1:(plus_group_age - 1)] * exp(-M)
  ## plus group
  N_next_year_age[,plus_group_age] = N_next_year_age[,plus_group_age] * exp(-M) + N_age[,plus_group_age] * exp(-M)
  ## movement
  N_age = t(movement_matrix) %*% N_next_year_age
}
iterative_N_age = N_age
plot(1:plus_group_age, N_age[3,], xlab = "Age", ylab= "Initial numbers", ylim = c(0,R0* 1.4), type = "l", lwd = 3)
lines(1:plus_group_age, N_age[2,], lwd = 3, col = "red", lty = 2)
lines(1:plus_group_age, N_age[1,], lwd = 3, col = "blue", lty = 2)
legend('topright', col = c("black","red","blue"), legend = c("Region 3", "Region 2", "Region 1"), lwd = 3)

```

```{r spatialcase_alt}
N_age = matrix(0, nrow = n_regions, ncol = plus_group_age)
update_N_age = N_age
for(i in 1:(plus_group_age)) {
  # recruitment
  update_N_age[,1] = R0 * exp(-M)
  # ageing and mortality
  update_N_age[,2:plus_group_age] = N_age[,1:(plus_group_age - 1)] * exp(-M)
  # plus group
  update_N_age[,plus_group_age] = update_N_age[,plus_group_age] * exp(-M) + N_age[,plus_group_age] * exp(-M)
  # movement
  N_age = t(movement_matrix) %*% update_N_age
}
## calculate one more annual cycle
# recruitment
update_N_age[,1] = R0 * exp(-M)
# ageing and mortality
update_N_age[,2:plus_group_age] = N_age[,1:(plus_group_age - 1)] * exp(-M)
# plus group
update_N_age[,plus_group_age] = update_N_age[,plus_group_age] * exp(-M) + N_age[,plus_group_age] * exp(-M)
# movement
update_N_age = t(movement_matrix) %*% update_N_age
## approximate!
c = update_N_age[,plus_group_age] / N_age[,plus_group_age] - 1
update_N_age[,plus_group_age] = N_age[,plus_group_age] * 1 / (1 - c)

iterative_N_age[,plus_group_age]
update_N_age[,plus_group_age]

plot(1:plus_group_age, iterative_N_age[3,], xlab = "Age", ylab= "Initial numbers", ylim = c(0,R0* 1.4), type = "l", lwd = 3)
lines(1:plus_group_age, update_N_age[3,], lwd = 3, col = "red", lty = 2)

lines(1:plus_group_age, iterative_N_age[2,], lwd = 3, col = "black", lty = 1)
lines(1:plus_group_age, update_N_age[2,], lwd = 3, col = "red", lty = 2)
lines(1:plus_group_age, iterative_N_age[1,], lwd = 3, col = "black", lty = 1)
lines(1:plus_group_age, update_N_age[1,], lwd = 3, col = "red", lty = 2)
legend('topright', col = c("black","red"), legend = c("Numerical", "Analytical"), lwd = 3, lty = c(1,2))

```



