# Estimating age-based movement {#agebasedmovement}

Simple tagging process model

\begin{align*}
T^k_{a, r,t} & = T_a^k, \quad t = 0\\
T^k_{a, r,t + 1} & = T^k_{a, r,t} e^{-M}, \quad t > 0\\
\boldsymbol{T}^k_{a, t + 1} & = \boldsymbol{T}^k_{a, t + 1} (1 - S^o_a)  \boldsymbol{M}^y + \boldsymbol{T}^k_{a, t + 1}  S^o_a \boldsymbol{M}^o
\end{align*}
where, \(T_a^k\) is the initial number of tag releases for release event \(k\) and age \(a\), \(\boldsymbol{T}^k_{a, t + 1}\) is a matrix of numbers of tagged fish across all regions and ages from release event \(k\) and \(\boldsymbol{M}^y\) is a \(n_r \times n_r\) movement matrix for young fish and \(\boldsymbol{M}^o\) is the movement matrix for older fish. \(S^o_a\) is a movement based logistic ogive that determines what proportion of age-cohorts move based on \(\boldsymbol{M}^o\) and \((1 - S^o_a)\) indicates what proportion of age-cohorts move based on \(\boldsymbol{M}^y\)



Generating age-based observation model 

\begin{align*}
\widehat{Y}^k_{r,t} & = \sum_a T^k_{a, r,t}(1 - S^o_a) \\
\widehat{O}^k_{r,t} & = \sum_a T^k_{a, r,t} S^o_a \\
{Y}^k_{r,t} & = \sum_a N^k_{a, r,t}(1 - S^o_a) \\
{O}^k_{r,t} & = \sum_a N^k_{a, r,t} S^o_a \\
\end{align*}

\(\widehat{Y}^k_{r,t}\) expected tag recoveries for young fish, and \(\widehat{O}^k_{r,t}\) is the expected tag recoveries for old fish, and \( N^k_{a, r,t}\) is the observed tag-recovery by age.


```{r loadagebasedTMBmodel, echo = F, eval = T, results = 'hide', warning=FALSE, message=FALSE, error = FALSE}
compile(file.path("TMB","SimpleAgeBasedMovement.cpp"), flags = "-Wignored-attributes -O3",DLLFLAGS="");
dyn.load(dynlib(file.path("TMB","SimpleAgeBasedMovement")))
```


```{r SetupAgeBasedOM}
M = 0.12
R0 = 20000
ages = 1:30
n_regions = 3
young_move_matrix = old_move_matrix = matrix(rnorm(n_regions * n_regions, 10, 3), nrow = n_regions, ncol = n_regions)
## young movement
## encouraged to stay closer to region 1
young_move_matrix[1,1] = 0.75
young_move_matrix[1,2] = 0.2
young_move_matrix[1,3] = 0.05

young_move_matrix[2,1] = 0.3
young_move_matrix[2,2] = 0.6
young_move_matrix[2,3] = 0.1

young_move_matrix[3,1] = 0.6
young_move_matrix[3,2] = 0.3
young_move_matrix[3,3] = 0.1

## old movement
## encouraged to stay closer to region 3
old_move_matrix[1,1] = 0.1
old_move_matrix[1,2] = 0.4
old_move_matrix[1,3] = 0.5

old_move_matrix[2,1] = 0.1
old_move_matrix[2,2] = 0.5
old_move_matrix[2,3] = 0.4

old_move_matrix[3,1] = 0.1
old_move_matrix[3,2] = 0.2
old_move_matrix[3,3] = 0.7

true_young_molten_move = melt(young_move_matrix)
true_old_molten_move = melt(old_move_matrix)

colnames(true_young_molten_move) = colnames(true_old_molten_move) = c("From", "To", "movement")

## age based movement ogive
a50_move = 6.7
a95_move = 1.7
old_age_based_movement = logis(ages, a50_move, a95_move)
young_age_based_movement = 1 - old_age_based_movement

init_N_age = calculate_initial_numbers_at_age_age_based_movement(n_regions, n_ages = length(ages), R0, old_movement_matrix = old_move_matrix, young_movement_matrix = young_move_matrix, age_based_movement_ogive = young_age_based_movement, natural_mortality =rep(M, length(ages)))

data = list()
data$initial_age_tag_releases = t(init_N_age)
data$natural_mortality = M
data$n_years = 5
data$n_ages = length(ages)
data$n_regions = n_regions
data$ages = ages
data$tag_recovery_obs = array(100, dim = c(data$n_ages, n_regions, n_regions, data$n_years))
data$a50_bounds = c(3, 15)
parameters = list()
parameters$logisitic_a50_movement = logit_general(a50_move, data$a50_bounds[1], data$a50_bounds[2])
parameters$ln_ato95_movement = log(a95_move)

parameters$transformed_movement_pars_young = array(0, dim = c(n_regions - 1, n_regions))
parameters$transformed_movement_pars_old = array(0, dim = c(n_regions - 1, n_regions))

for(r in 1:n_regions) {
  parameters$transformed_movement_pars_young[,r] = simplex(young_move_matrix[r, ])
  parameters$transformed_movement_pars_old[,r] = simplex(old_move_matrix[r, ])
}
test_obj <- MakeADFun(data = data, parameters = parameters, DLL="SimpleAgeBasedMovement")

```


## Simulation with fixed age-based ogive {-}
Run a simulation with age-based movement ogive is fixed at correct values

```{r FixOgiveSim}
###### set starting parameters
start_movement_matrix = matrix(rlnorm(n = n_regions * n_regions, log(0.01), 0.1), n_regions, n_regions)
# renormalise
start_movement_matrix = sweep(start_movement_matrix, 1, STATS = rowSums(start_movement_matrix), "/")

start_parameters = list()
start_parameters$logisitic_a50_movement = logit_general(4, data$a50_bounds[1], data$a50_bounds[2])
start_parameters$ln_ato95_movement = log(1)

start_parameters$transformed_movement_pars_young = array(0, dim = c(n_regions - 1, n_regions))
start_parameters$transformed_movement_pars_old = array(0, dim = c(n_regions - 1, n_regions))

for(r in 1:n_regions) {
  start_parameters$transformed_movement_pars_young[,r] = simplex(start_movement_matrix[r, ])
  start_parameters$transformed_movement_pars_old[,r] = simplex(start_movement_matrix[r, ])
}

n_sims = 250;
sim_ls = list();
young_age_based_ogive_df = old_age_based_ogive_df = NULL
young_matrix_df = old_matrix_df = NULL
a50_vec = ato95_vec = NULL
## fix selectivity parameters at true values
na_map = fix_pars(par_list = start_parameters, pars_to_exclude = c("logisitic_a50_movement","ln_ato95_movement"))
# fix at true values
start_parameters$logisitic_a50_movement = logit_general(a50_move, data$a50_bounds[1], data$a50_bounds[2])
start_parameters$ln_ato95_movement = log(a95_move)
nll = vector()
for(sim_iter in 1:n_sims){
  ## simulate data
  simdata = test_obj$simulate(complete = T)
  
  ## re-estimate
  reest_obj <- MakeADFun(data = simdata, parameters = start_parameters, map = na_map, DLL="Simple",
                         silent = T)
  #reest_obj$fn()
  #reest_obj$gr()
  ## Estimate
  mle_spatial = nlminb(start = reest_obj$par, objective = reest_obj$fn, gradient  = reest_obj$gr,
                       control = list(iter.max = 10000, eval.max = 10000))
  try_improve = tryCatch(expr =
                           for(i in 1:2) {
                             g = as.numeric(reest_obj$gr(mle_spatial$par))
                             h = optimHess(mle_spatial$par, fn = reest_obj$fn, gr = reest_obj$gr)
                             mle_spatial$par = mle_spatial$par - solve(h,g)
                             mle_spatial$objective = reest_obj$fn(mle_spatial$par)
                           }
                         , error = function(e){e})
  
  try_improve
  
  mle_report = reest_obj$report(mle_spatial$par)
  
  a50_vec = c(a50_vec, mle_report$a50_movement)
  ato95_vec = c(ato95_vec, mle_report$ato95_movement)
  
  
  young_ogive = data.frame(age = data$ages, ogive = mle_report$young_age_based_movement_ogive, sim = sim_iter)
  old_ogive = data.frame(age = data$ages, ogive = mle_report$old_age_based_movement_ogive, sim = sim_iter)
  
  young_age_based_ogive_df = rbind(young_age_based_ogive_df, young_ogive)
  old_age_based_ogive_df = rbind(old_age_based_ogive_df, old_ogive)
  
  young_matrix = mle_report$movement_matrix_young
  old_matrix = mle_report$movement_matrix_old
  
  young_molten_move = melt(young_matrix)
  old_molten_move = melt(old_matrix)
  
  colnames(young_molten_move) = colnames(old_molten_move) = c("From", "To", "movement")
  young_molten_move$sim = sim_iter
  old_molten_move$sim = sim_iter
  
  young_matrix_df = rbind(young_matrix_df, young_molten_move)
  old_matrix_df = rbind(old_matrix_df, old_molten_move)
  
  nll = c(nll, mle_spatial$objective)
}


## visualise young movement rates
ggplot(young_matrix_df) +
  geom_boxplot(aes(y = movement)) +
  geom_hline(data = true_young_molten_move, aes(yintercept = movement, col = "OM"), 
             linewidth = 1.1, linetype = "dashed") +
  facet_grid(From~To) +
  ggtitle("Young movement rates") +
  ylim(0,1) +
  theme_bw()


ggplot(old_matrix_df) +
  geom_boxplot(aes(y = movement)) +
  geom_hline(data = true_old_molten_move, aes(yintercept = movement, col = "OM"), 
             linewidth = 1.1, linetype = "dashed") +
  facet_grid(From~To)  +
  ggtitle("old movement rates") +
  ylim(0,1) +
  theme_bw()
```



## Simulation estimate a50 parameter of age-based ogive {-}


```{r Esta50Ogive}
###### set starting parameters
sim_ls = list();
young_age_based_ogive_df_a50 = old_age_based_ogive_df_a50 = NULL
young_matrix_df_a50 = old_matrix_df_a50 = NULL
a50_vec = ato95_vec = NULL
## fix selectivity parameters at true values
na_map = fix_pars(par_list = start_parameters, pars_to_exclude = c("ln_ato95_movement"))
# fix at true values
start_parameters$logisitic_a50_movement = logit_general(4, data$a50_bounds[1], data$a50_bounds[2])
start_parameters$ln_ato95_movement = log(a95_move)
nll_a50 = vector()
for(sim_iter in 1:n_sims){
  ## simulate data
  simdata = test_obj$simulate(complete = T)
  
  ## re-estimate
  reest_obj <- MakeADFun(data = simdata, parameters = start_parameters, map = na_map, DLL="Simple", silent = T)
  #reest_obj$fn()
  #reest_obj$gr()
  ## Estimate
  mle_spatial = nlminb(start = reest_obj$par, objective = reest_obj$fn, gradient  = reest_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
  try_improve = tryCatch(expr =
                           for(i in 1:2) {
                             g = as.numeric(reest_obj$gr(mle_spatial$par))
                             h = optimHess(mle_spatial$par, fn = reest_obj$fn, gr = reest_obj$gr)
                             mle_spatial$par = mle_spatial$par - solve(h,g)
                             mle_spatial$objective = reest_obj$fn(mle_spatial$par)
                           }
                         , error = function(e){e})
  
  try_improve
  
  mle_report = reest_obj$report(mle_spatial$par)
  
  a50_vec = c(a50_vec, mle_report$a50_movement)
  ato95_vec = c(ato95_vec, mle_report$ato95_movement)
  
  
  young_ogive = data.frame(age = data$ages, ogive = mle_report$young_age_based_movement_ogive, sim = sim_iter)
  old_ogive = data.frame(age = data$ages, ogive = mle_report$old_age_based_movement_ogive, sim = sim_iter)
  
  young_age_based_ogive_df_a50 = rbind(young_age_based_ogive_df_a50, young_ogive)
  old_age_based_ogive_df_a50 = rbind(old_age_based_ogive_df_a50, old_ogive)
  
  young_matrix = mle_report$movement_matrix_young
  old_matrix = mle_report$movement_matrix_old
  
  young_molten_move = melt(young_matrix)
  old_molten_move = melt(old_matrix)
  
  colnames(young_molten_move) = colnames(old_molten_move) = c("From", "To", "movement")
  young_molten_move$sim = sim_iter
  old_molten_move$sim = sim_iter
  
  young_matrix_df_a50 = rbind(young_matrix_df_a50, young_molten_move)
  old_matrix_df_a50 = rbind(old_matrix_df_a50, old_molten_move)
  
  nll_a50 = c(nll_a50, mle_spatial$objective)
}

ggplot(young_matrix_df_a50) +
  geom_boxplot(aes(y = movement)) +
  geom_hline(data = true_young_molten_move, aes(yintercept = movement, col = "OM"), linewidth = 1.1, linetype = "dashed") +
  facet_grid(From~To) +
  ggtitle("Young movement rates") +
  ylim(0,1) +
  theme_bw()


ggplot(old_matrix_df_a50) +
  geom_boxplot(aes(y = movement)) +
  geom_hline(data = true_old_molten_move, aes(yintercept = movement, col = "OM"), linewidth = 1.1, linetype = "dashed") +
  facet_grid(From~To)  +
  ggtitle("Old movement rates") +

  ylim(0,1) +
  theme_bw()

hist(a50_vec)
abline(v = a50_move, col = "red", lty = 2, lwd = 3)

```




