# Tag likelihoods {#taglike}
This section explores a range of tag-recovery likelihoods that have been used in the literature. I simple tag recapture model is used in a simulation to explore how well different likelihoods estimate movement parameters under a range of dispersion conditions. The tag-recapture model ignores age and models number of tagged fish from a release event denoted by \(k\) over three regions indexed by \(r\) for \(n_t\) discrete time-steps. \(T^k_{r,t}\) denotes the number of tagged fish from release event \(k\) in region \(r\) and time step \(t\). The model applies the following processes between time-steps.

\begin{align*}
T^k_{r,t} & = T^k, \quad t = 0\\
T^k_{r,t + 1} & = T^k_{r,t} e^{-Z_t}, \quad t > 0\\
Z_t & = M + F_t \\
\boldsymbol{T}^k_{t + 1} & = \boldsymbol{T}^k_{t + 1} \boldsymbol{M}
\end{align*}
where, \(T^k\) is the initial nunmber of tag releases for release event \(k\), \(\boldsymbol{T}^k_{t + 1} =  (T^k_{1,t + 1},\dots, T^k_{n_r,t + 1})\) is a vector of numbers of tagged fish across all regions and \(\boldsymbol{M}\) is a \(n_r \times n_r\) movement matrix. This assumes each region has the same fishing mortality in each time-step.

## Likelihoods {-}

For all possible tag-recovery events indexed by \(m\) (where \(m = \{r,t\}\) i.e., has an implied region and time-step dimension) the tag-recapture model derives expected tag-recoveries. In this case, any time or region that has any fishing mortality rate is considered a possible tag-recovery event. All the likelihoods in the following sections use the same derivation for expected tag-recoveries, which follows

\[
N^k_{m} = T^k_{r,t} \left(1 - e^{-Z_t}\right) \frac{F_t}{Z_t}, \quad m = \{r,t\} \ .
\]


### Poisson {-}
The Poisson is the simplest likelihood, which was explored by @hilborn1990determination. It assumes

\[
N^k_{m} \sim \mathcal{Poisson}(\widehat{N}^k_{m}) \,
\]

with the total log-likelihood contribution evaluating the Poisson log-likelihood over all possible release and recovery events (\(\sum\limits_k\sum\limits_m \))

### Negative Binomial {-}
A similar likelihood to the Poisson is the Negative Binomial which was explored by @hanselman2015move. This is a more flexible likelihood as it allows for over dispersion, through an additional estimable parameter denoted by \(\phi\)

\[
N^k_{m} \sim \mathcal{NB}(\widehat{N}^k_{m}, \widehat{\phi}) \,
\]

with the total log-likelihood contribution evaluating the Negative binomial log-likelihood over all possible release and recovery events (\(\sum\limits_k\sum\limits_m \))


### Multinomial:release conditioned {-}

The release conditioned multinomial likelihood was explored by @polacheck2006integrating and subsequently by @vincent2020parameter \& @goethel2014demonstration. This treats all recovery events plus an additional not-recovered event as a multinomial distributed event.

\[
\boldsymbol{N}^k \sim \mathcal{Multinomial}\left(\widehat{\boldsymbol{\theta}}^k\right)
\]
where, \(\boldsymbol{N}^k\) is the vector of tag-recoveries for all possible recovery events (\(n_m\)) plus an additional not-recovered event.
\[
\boldsymbol{N}^k = (N^k_1, N^k_2, \dots, N^k_{n_m}, N^k_{NR})
\]
where, \(N^k_{NR}\) represents the not recovered category and is \(N^k_{NR} = T^k - \sum\limits_m^{n_m} N^k_m\), where \(T^k\) is the initial number of tags released for event \(k\). \(\widehat{\boldsymbol{\theta}}^k\) is a vector of proportions that sum to one, with the same dimensions as \(\boldsymbol{N}^k\) and is calculated as,

\[
\widehat{\theta}^k_m = \frac{\widehat{N}^k_{m}}{T^k}, \quad \forall \ m \in (1, \dots,n_m)
\]
where, again \(T^k\) is the initial number of tags released for event \(k\). The proportions for the not recaptured group is calculated as,
\[
\widehat{\theta}^k_{NR} = 1 - \sum_m^{n_m} \widehat{\theta}^k_m 
\]


### Recapture conditioned {-}

The multinomial which is recapture conditioned was described by @vincent2020parameter but based on the original work of @mcgarvey2002estimating (described in the next section). This likelihood evaluates the probability of recapturing a tagged fish in a certain region among all possible regions for a given release event and time-period.


\[
\boldsymbol{N}^k_t \sim \mathcal{Multinomial}\left(\widehat{\boldsymbol{\theta}}_t^k\right)
\]
where, \(\boldsymbol{N}^k_t\) is the vector of tag-recoveries in all regions (\(n_m\)) for release event \(k\) and time period \(t\)
\[
\boldsymbol{N}^k_t = (N^k_{1,t}, N^k_{2,t}, \dots, N^k_{n_r,t})
\]
and model predicted proportions
\[
\widehat{\boldsymbol{\theta}}_t^k = (\widehat{\theta}^k_{1,t}, \widehat{\theta}^k_{2,t}, \dots, \widehat{\theta}^k_{n_r,t})
\]
and,

\[
\widehat{\theta}^k_{r,t} = \frac{\widehat{\theta}^k_{r,t}}{\sum\limits_r \widehat{\theta}^k_{r,t}}
\]


The literature [@mcgarvey2002estimating;@vincent2020parameter] often describes the log likelihood as

\[
ll = \sum_k\sum_y\sum_r log(\widehat{\theta}^k_{r,t}) \times N^k_{r,t}
\]



## Simple simulation {-}
To explore these likelihoods are very simple simulation was setup to see how well each likelihood could estimate elements of \(\boldsymbol{M}\) under different levels of over dispersion. The simulation assumes all tags encountered where reported (100\% reporting rate) and both natural mortality and fishing mortality are known without error. The only estimated parameters were \(\boldsymbol{M}\) and \(\phi\) when the negative binomial likelihood was explored.


The following is some R-code that was used to generate operating model values for 
```{r simulate_tag_data}
set.seed(123)
n_y = 5
n_regions = 3
M = 0.13
F_y = rlnorm(n = n_y, log(0.2), 0.6)
Z_y = F_y + M
S_y = exp(-Z_y)
# make up some random movement matrix
move_matrix = matrix(0, ncol = n_regions, nrow = n_regions)
diag(move_matrix) = c(0.7, 0.5, 0.6)
move_matrix = move_matrix + rnorm(n = n_regions * n_regions, 0.1,0.05)
move_matrix = sweep(move_matrix, MARGIN = 1, STATS = rowSums(move_matrix), FUN = "/")
## seed tag releases
tag_release_by_region = rep(1000, n_regions)
n_release_events = n_regions
## Calculate tag-partition
tags_partition_by_release_event = array(0, dim = c(n_regions, n_y + 1, n_release_events))
expected_tag_recoveries  = tag_recovery_obs = array(0, dim = c(n_regions, n_y, n_release_events))
for(r in 1:n_release_events)
  tags_partition_by_release_event[r,1,r] = tag_release_by_region[r]
for(y in 1:n_y) {
  for(rel_event in 1:n_release_events) {
    ## ageing and F
    tags_partition_by_release_event[,y + 1, rel_event] = 
      tags_partition_by_release_event[,y,rel_event] * S_y[y]
    ## Movement
    tags_partition_by_release_event[,y + 1, rel_event] = 
      tags_partition_by_release_event[,y + 1, rel_event] %*% move_matrix
  }
}

## Create tag-recovery expected values
for(rel_event in 1:n_regions) {
  expected_tag_recoveries[,,rel_event] = 
    sweep(tags_partition_by_release_event[ ,2:(n_y + 1),rel_event], MARGIN = 2, STATS =  F_y / Z_y * (1 - S_y), FUN = "*")
}
```


The following code chunk assumes `tags_partition_by_release_event` is the same among simulations, but uses the negative binomial with a range of over-dispersion parameters to simulate tag-recoveries for `n_sim` times. We explore Relative error (RE) and mean square error (MSE) for the different likelihood choices.

```{r loadtagTMBmodels, echo = F, eval = T, results = 'hide', warning=FALSE, message=FALSE, error = FALSE}
compile(file.path("TMB","SimpleTagEstimator.cpp"), flags = "-Wignored-attributes -O3",DLLFLAGS="");
dyn.load(dynlib(file.path("TMB","SimpleTagEstimator")))
```

```{r OM_EM}
n_sim = 500
## Populate TMB objects
data = list()
data$n_y = n_y
data$n_regions = n_regions
data$M = M
data$F_y = F_y
data$tag_likelihood = 0 ## Poisson
data$tag_recovery_obs = tag_recovery_obs
data$number_of_tags_released = tag_release_by_region
data$movement_transformation = 0 ## simplex, can use the logistic transformation as well
## starting values
est_parameters = list()
start_matrix = matrix(0, ncol = n_regions, nrow = n_regions)
diag(start_matrix) = c(0.95, 0.95, 0.95)
start_matrix = start_matrix + rnorm(n = n_regions * n_regions, 0.02,0.0001)
## force to sum = 1
start_matrix = sweep(start_matrix, MARGIN = 1, STATS = rowSums(start_matrix), FUN = "/")

est_parameters$transformed_movement_pars = matrix(NA, nrow = n_regions - 1, ncol = n_regions)
for(i in 1:n_regions)
  est_parameters$transformed_movement_pars[,i] = simplex(start_matrix[i,])
est_parameters$ln_phi = log(1)
## don't estimate phi parameter unless Negative binomial
na_map = fix_pars(par_list = est_parameters, pars_to_exclude = "ln_phi")

OM_mat = melt(move_matrix)
colnames(OM_mat) = c("From", "To","proportion")
OM_mat$type = "OM"
```



```{r runsimulationTagLike, warning=FALSE, message=FALSE, error = FALSE, cache=TRUE}
############################################
## over_diserpsion_investication
## what happens when we simualte from
## negative binomial with different 
## over-dispersion parameters
############################################

over_dispersion_params = c(0.1, 0.2, 0.3, 0.5, 1, 10)
n_sim = 500
# use the simplex
data$movement_transformation = 0
sim_data_info = NULL
complete_df = NULL
for(over_ndx in 1:length(over_dispersion_params)) {
  this_dispersion = over_dispersion_params[over_ndx]
  full_move_df = NULL;
  temp_proportion_zeros_sim_data = temp_variance_sim_data = c()
  for(sim in 1:n_sim) {
    ## keep expected values the same among simulations
    for(rel_event in 1:n_release_events) {
      for(y in 1:n_y) {
        for(r in 1:n_regions) {
          tag_recovery_obs[r,y,rel_event] = rnbinom(n = 1, mu = expected_tag_recoveries[ r,y,rel_event], size = this_dispersion)
        }
      }
    }
    ## summarise number of zeros and variance of observed data
    temp_proportion_zeros_sim_data = c(temp_proportion_zeros_sim_data, sum(tag_recovery_obs == 0));
    temp_variance_sim_data = c(temp_variance_sim_data, var(tag_recovery_obs));
    ## save data
    data$tag_recovery_obs = tag_recovery_obs

    ## Poisson
    data$tag_likelihood = 0
    est_obj <- MakeADFun(data, est_parameters, map = na_map, DLL="SimpleTagEstimator", silent = T)
    est_obj$env$tracepar = F
    opt_poisson = nlminb(est_obj$par, est_obj$fn, est_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
    poisson_rep = est_obj$report(opt_poisson$par)
    sd_poisson = sdreport(est_obj)

    ## Multinomial release conditioned
    data$tag_likelihood = 1
    est_obj <- MakeADFun(data, est_parameters, map = na_map, DLL="SimpleTagEstimator", silent =T)
    opt_multi_release = nlminb(est_obj$par, est_obj$fn, est_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
    multi_release_rep = est_obj$report(opt_multi_release$par)
    sd_multi_release = sdreport(est_obj)

    ## Multinomial recapture conditioned
    data$tag_likelihood = 2
    est_obj <- MakeADFun(data, est_parameters,map = na_map,  DLL="SimpleTagEstimator", silent = T)
    opt_multi_recap = nlminb(est_obj$par, est_obj$fn, est_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
    multi_recap_rep = est_obj$report(opt_multi_recap$par)
    sd_multi_recap = sdreport(est_obj)

    ## recapture conditioned
    data$tag_likelihood = 3
    est_obj <- MakeADFun(data, est_parameters,map = na_map,  DLL="SimpleTagEstimator", silent = T)
    opt_recap = nlminb(est_obj$par, est_obj$fn, est_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
    recap_rep = est_obj$report(opt_recap$par)
    sd_recap = sdreport(est_obj)
    ## Negative binomial
    data$tag_likelihood = 4
    est_obj <- MakeADFun(data, est_parameters,  DLL="SimpleTagEstimator", silent = T)
    opt_nb = nlminb(est_obj$par, est_obj$fn, est_obj$gr, control = list(iter.max = 10000, eval.max = 10000))
    nb_rep = est_obj$report(opt_nb$par)
    sd_nb = sdreport(est_obj)
    ## reformat movement estimates
    pois_mat = melt(poisson_rep$movement_matrix)
    pois_mat$RE = (pois_mat$value - OM_mat$proportion) / OM_mat$proportion * 100
    pois_mat$SE = (OM_mat$proportion - pois_mat$value)^2
    pois_mat$type = "Poisson"
    pois_mat$sim = sim
    recap_mat = melt(recap_rep$movement_matrix)
    recap_mat$RE = (recap_mat$value - OM_mat$proportion) / OM_mat$proportion * 100
    recap_mat$SE = (OM_mat$proportion - recap_mat$value)^2
    recap_mat$type = "Recapture"
    recap_mat$sim = sim
    multi_release_mat = melt(multi_release_rep$movement_matrix)
    multi_release_mat$RE = (multi_release_mat$value - OM_mat$proportion) / OM_mat$proportion * 100
    multi_release_mat$SE = (OM_mat$proportion - multi_release_mat$value)^2
    multi_release_mat$type = "Multinomial Release"
    multi_release_mat$sim = sim
    nb_mat = melt(nb_rep$movement_matrix)
    nb_mat$RE = (nb_mat$value - OM_mat$proportion) / OM_mat$proportion * 100
    nb_mat$SE = (OM_mat$proportion - nb_mat$value)^2
    nb_mat$type = "Negative Binomial"
    nb_mat$sim = sim
    full_move_df = rbind(full_move_df, rbind(pois_mat, recap_mat, multi_release_mat, nb_mat))
  }
  prop_obs_zero = temp_proportion_zeros_sim_data / (dim(data$tag_recovery_obs)[1] * dim(data$tag_recovery_obs)[2] * dim(data$tag_recovery_obs)[3])
  tmp_df = data.frame(avg_proportion_zero = mean(prop_obs_zero), avg_var = mean(temp_variance_sim_data), over_dispersion_param = this_dispersion)
  sim_data_info = rbind(sim_data_info, tmp_df)
  full_move_df$over_dispersion_param = this_dispersion
  complete_df = rbind(complete_df, full_move_df)
}
colnames(complete_df) = c("From", "To","proportion", "RE", "SE", "type", "sim", "over_dispersion_param")
```


```{r overdispersiontable, echo = F, eval = T, warning=FALSE, message=FALSE, error = FALSE}
kable(sim_data_info, caption = "Average proportion of zeros and variance of simulated data for different overdispersion settings.", digits  = 3)
```


```{r taglikesummaryplotsRE, echo = F, eval = T, warning=FALSE, message=FALSE, error = FALSE, fig.cap="Relative error in movement parameters. Columns are overdispersion values for simulated data. GGplot facets are movement matrix elemen combinatons", fig.height=10}
ggplot(complete_df, aes(y = RE, col = type, x = type)) +
  geom_boxplot() +
  ylim(-100, 150) +
  labs(x = "", y= "Relative error (%)", col = "Likelihood") +
  geom_hline(yintercept = 0, col = "red", linetype = "dashed") +
  theme_bw() +
  facet_grid(From~To~over_dispersion_param) +
  theme(axis.text.x = element_blank())
```


```{r taglikesummaryplotsEst, echo = F, eval = T, warning=FALSE, message=FALSE, error = FALSE, fig.cap="Estimated movement parameters, red line is true value. Columns are overdispersion values for simulated data. GGplot facets are movement matrix elemen combinatons", fig.height=10}

ggplot(complete_df) +
  geom_boxplot(aes(y = proportion, col = type, x = type)) +
  geom_hline(data =OM_mat, aes(yintercept = proportion), col = "red", linetype = "dashed", linewidth = 1.1) +
  ylim(0, 1) +
  labs(x = "", y= "Estimated Movement parameters", col = "Likelihood") +
  theme_bw() +
  facet_grid(From~To~over_dispersion_param) +
  theme(axis.text.x = element_blank())


```


```{r taglikesummaryplotsMSE, echo = F, eval = T, warning=FALSE, message=FALSE, error = FALSE, fig.cap="Mean squared error (MSE) over all simulations and estimated parameters."}

ggplot(complete_df %>% group_by(type, over_dispersion_param) %>% summarise(MSE = sum(SE) / n_sim)) +
  geom_col(aes(y = MSE, col = type, fill = type, x= type)) +
  facet_wrap(~over_dispersion_param) +
  theme_bw() +
  labs(x = "", col = "Likelihood", fill = "Likelihood", y = "MSE") +
  theme(axis.text.x = element_blank())
```